---
title: "ESM_22_HW3"
author: "Camila Bobroff"
date: "3/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, inlcude=FALSE, message=FALSE, warning=FALSE}
#Part 0. Load packages
library(tidyverse)
library(tseries)
library(forecast)
```

####Task 1. Open science perspectives

#####600 - 800 word statement

####Task 2. Truckee River flow (2000 - 2016)

#####a) Graph with the decomposed time series information
```{r, message=FALSE, warning=FALSE}
# Get data
truckee <- read_csv("clean_truckee_flow.csv")

# Convert to ts data
truckee_ts <- ts(truckee$mean_va, frequency = 12, start = c(2000,1))
# plot(truckee_ts)

# Decompose to explore data further
truckee_dc <- decompose(truckee_ts)
plot(truckee_dc)
```
#####Although there are some high values and low values in the data, there do not seem to be any outliers (which could have a disproportionate effect on the time series model). There seems to be a slight downward trend in the data and it looks non-stationary and additive. We see a seasonal pattern that repeats about every 12 months and a slight cyclical trend about every five years (the seasonal component is on the same scale as the orinigal data).

#####b) Forecast the Truckee River for 5 years after the final observation in the dataset
```{r, message=FALSE, warning=FALSE}
# Holt Winters exponential smoothing
truckee_hw <- HoltWinters(truckee_ts)
# plot(truckee_hw)

# Forecast Holt Winters
truckee_forecast <- forecast(truckee_hw, h = 60)
# plot(truckee_forecast)

# Autoregressive integrated moving average (ARIMA) for comparison
# estimate pdq

truckee_pdq <- auto.arima(truckee_ts) # [2,1,1,][0,0,2]

# fit the ARIMA model
truckee_arima <- arima(truckee_ts, order = c(2,1,1), seasonal = list(order = c(0,0,2)))

# evaluate residuals
# par(mfrow = c(1,2))
# hist(truckee_arima$residuals)
# qqnorm(truckee_arima$residuals) # looks normal

# forecast ARIMA
forecast_truckee <- forecast(truckee_arima, h = 60)
# plot(forecast_truckee) 

# Graph of Holt Winters 
plot(truckee_forecast,
     xlab = "Time",
     ylab = "Truckee River Flows (cubic feet per second)")

```

#####c) Holt Winters residuals
```{r, message=FALSE, warning=FALSE}
par(mfrow = c(1,2))
hist(truckee_forecast$residuals) 
qqnorm(truckee_forecast$residuals) # Looks relatively normally distributed.
```

